Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-05 tex/main.tex
more edits

If $t$ has positive polarity, we construct a conjunction of the result of the
recursive calls with the formulas in $D = D_1 \mathrel\cup \cdots \mathrel\cup
D_n$. Similarly, if $t$ has negative polarity, we construct a disjunction of
the result of the recursive calls with the negation of formulas $D_1, \ldots,
D_n$. It $t$ has no polarity, then we return the union of $D_1 \cup \ldots \cup
D_n$.

If $t$ is a quantified formula $\forallf{\con{f}} \vec x.\; t_1$, it
recursively calls $\conv$ on its body with the same polarity. If $t$ is a
function definition of a function in $\sfundefs{\Sigma}$, we instead construct
a quantified formula over a single variable $y$ of type $\fargsort{\con{f}}$,
and replace all occurrences of $\vec x$ in $t'_1$ with $\vecfarg{\con{f}}( y )$.
In this case, since function definitions are a top-level conjunct, by case
analysis on the return values of $\conv$, we know that $D_1$ is empty. In the
case that $t$ is not a function definition, the resulting quantified formula is
reconstructed with the term $t'_1$ returned by the recursive call, and a
quantified prefix is appended to the formula $D_1$ returned by the recursive
call.

Otherwise, $t$ is either an application of an uninterpreted predicate, or a
term of a type other than $\Bool$. In this case, the second component of our
return value contains a constraint of the form $\exists
z\typ{\fargsort{\con{f}}}. ( \vecfarg{\con{f}}( z ) \teq \vec t )$ for each
subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}$ is a function in
$\sfundefs{\Sigma}$. When such constraints are asserted positively, this
indicates that the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must
include $\vec t$ in its range.

We demonstrate this translation with an example.

>>>>

If $t$ has positive polarity, $\conv_0$ adds the constraint $\chi$ derived from
the subterms as a conjunction. Dually, if $t$ has negative polarity, the
function constructs a disjunction with the negated constraint. It $t$ has no
known polarity, the constrained $\chi$ is simply returned to the caller.

If $t$ is a quantified formula, $\conv_0$ recursively calls itself on the body
with the same polarity. If $t$ is a function definition of a function in
$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
$y$ of type $\fargsort{\con{f}}$ and replaces all occurrences of $\vec x$ in
$t'_1$ with $\vecfarg{\con{f}}( y )$.Since function definitions
are a top-level conjunct, %by case analysis on the return values of $\conv$,
we know that $\chi$ is simply $\true$.
If $t$ is not a function definition, the resulting
quantified formula is reconstructed with the body $u'$ returned by the
recursive call, and a quantifier is prefixed to the constraint $\chi$
returned by the call.

Otherwise, $t$ is either an application of an uninterpreted predicate or a term
of a type other than $\Bool$. The returned constraint is a conjunction of
formulas of the form $\exists z\typ{\fargsort{\con{f}}}.\;\, \vecfarg{\con{f}}(
z ) \teq \vec t$ for each subterm $\con{f}( \vec t )$ of $t$ such that $\con{f}
\in \sfundefs{\Sigma}$. Such constraints, when asserted positively, ensure that
the tuple of uninterpreted functions $\vecfarg{\con{f}}$ must include $\vec t$
in its range.

%We demonstrate this translation with an example.
